---
title: "RR Project 2"
output: html_document
date: "2023-06-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Processing

First we download the data as a compressed csv file using the bz2 compression and load it into the globale environment. This code block and following code blocks in the data processing section will cache the data to improve the processing. 

```{r Loading Data, cache=TRUE}
# Download and load csv file into Environment
library(dplyr, warn.conflicts = FALSE)
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2','./StormData.csv.bz2')
rdt <- read.csv('StormData.csv.bz2')
```

Next we will subset the data into only the portion that we need and check for classification errors in the event types that will signifigantly impact our ability to provide a complete analysis of the data.
```{r Data Subset and Cleaning, cache= TRUE}
subset_dt <- rdt %>% select(STATE, EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP, FATALITIES, INJURIES)
subset_dt <- subset_dt %>% mutate(EVTYPE = trimws(EVTYPE))
subset_dt <- subset_dt %>% mutate(EVTYPE = as.factor(tolower(EVTYPE)))
subset_dt <- subset_dt %>% mutate(EVTYPE = gsub("tstm wind","thunderstorm wind",EVTYPE))

# True_et data classifications taken from the NSW publication on the preperation of data.
True_et <- tibble( EVTYPE = c("astronomical low tide", "avalanche", "blizzard", "coastal flood", "cold/wind chill", "debris flow","dense fog",'dense smoke', "drought", "dust devil", "dust storm", "excessive heat", "extreme cold/wind chill", "flash flood", "flood", "freezing fog", "frost/freeze", "funnel cloud","hail", "heat", "heavy rain", "heavy snow", "high surf", "high wind", "hurricane/typhoon", "ice storm", "lakeshore flood", "lake-effect snow", "lightning", "marine hail", "marine high wind", "marine strong wind", "marine thunderstorm wind", "rip current", "seiche", "sleet", "storm tide","strong wind", "thunderstorm wind", "tornado", "tropical depression", "tropical storm","tsunami", "volcanic ash", "waterspout", "wildfire", "winter storm", "winter weather"))

working_dt <- inner_join(True_et,subset_dt, by = "EVTYPE")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
